---
title: "iroot"
author: "chiteri"
date: "7/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```

```{r}
#libraries
library(tidyverse)
```

```{r}
#read data
blups <- read.csv("./blups/blups_combined.csv")

#filter separate days
day12 <- blups %>%  filter(Day == 12)  #remove CMM, ImageFlag because they have no variance
day15 <- blups %>%  filter(Day == 15) 
day18 <- blups %>%  filter(Day == 18)

```

############################################################################################
##################################### Day 18 ################################################
############################################################################################

```{r}
############################################################################################
#create a dataframe for ranking of each PI under each trait

#take only the columns with numerical data
colnum = c(5:(ncol(day18))) # 

day18_ranks <- data.frame(matrix(vector(),223,1, dimnames=list(c(), c("PI")))) #create new dataframe that has 375 rows
day18_ranks$PI <- (day18$PI) #add column of PI numbers from dataframe

#colnames(day18)

i = 1
for (i in 1:length(colnum)) {  #this second loop runs through each TRAIT, one at a time
  x=colnum[i]  #set the current [i] column as x
  trait=colnames(day18)[x] #sets the current column header as the trait name
  #Rank <- as.data.frame(rank(desc(day18[x]))) #this was the original code. Needs to be relooked at...why desc. When merging, does it align with original PI/entry???
  Rank <- as.data.frame(rank(-(day18[x]))) #- means ranking from largest to smallest

    #Rank <- as.data.frame(rank(day18[x]))
  names(Rank)[1] <- trait  #set current trait as column header
  day18_ranks <- cbind(day18_ranks,Rank) #add columns to existing dataframe   
}

```
## iroot categories according to literature by Falk et al 2020
## Each category is associated with certain traits
## will only focus on topsoil foraging and steep, deep and cheap

```{r}
############################################################################################
##################################### Scavenger ############################################
############################################################################################

IdeotypeVar <-  day18_ranks %>%
                            select(TRLUpper,WID, TRL_GR)
#select RSA traits to use for particular ideotype
########################


IdeotypeSums <- IdeotypeVar %>%
                            mutate(SUM = IdeotypeVar %>% rowSums()) #sum the rankings to make a master ranking column
Meta <- day18_ranks %>%
                  select(PI) #Create meta df with PI 

Scavenger <- cbind(Meta, IdeotypeSums) #bind meta and ideotype df together

#add country
country_1 <- day18[,c(1,3)] 
Scavenger <- merge(country_1, Scavenger, by="PI")
  
Scavenger <- arrange(Scavenger,SUM)  %>%
mutate(ScavengerRanked = rank(SUM, ties.method = "max")) #sort new df based on master ranking column with no.1 being best



ScavengerTop10 <- as.data.frame(Scavenger[1:10,1:2, drop = FALSE])
ScavengerBottom10 <- as.data.frame(Scavenger[213:223,1:2, drop = FALSE])

#colnames(ScavengerTop10) <-  "Scavenger"
```


```{r}
############################################################################################
################################## Steep, Deep, Cheap#######################################
############################################################################################

day18_ranks <- mutate(day18_ranks, LRADescending = 224 - LRA)  #invert the ranking for LRAMediangle as a lower angle, not higher angle is desirable

day18_ranks <- mutate(day18_ranks, SOL2 = -SOL2)  #invert the ranking for SOL2 as a lower density, not higher density is desirable

IdeotypeVar <-  day18_ranks %>%
                            select(LRADescending,PRL,SOL2,TRL_GR) #select RSA traits to use for particular ideotype
IdeotypeSums <- IdeotypeVar %>%
                              mutate(SUM = IdeotypeVar %>% rowSums()) #sum the rankings to make a master ranking column
Meta <- day18_ranks %>%
                   select(PI) #Create meta df with Entry, PI and Cluster info

SteepDeepCheap <- cbind(Meta, IdeotypeSums) #bind meta and ideotype df together

#add country
country_1 <- day18[,c(1,3)] 
SteepDeepCheap <- merge(country_1, SteepDeepCheap, by="PI")
 
SteepDeepCheap <- arrange(SteepDeepCheap, SUM) %>% mutate(DTRanked = rank(SUM, ties.method = "max")) #sort new df based on master ranking column with no.1 being best

SteepDeepCheapTop10 <- as.data.frame(SteepDeepCheap[1:10,1:2, drop = FALSE])
SteepDeepCheapBottom10 <- as.data.frame(SteepDeepCheap[213:223,1:2, drop = FALSE])
#colnames(SteepDeepCheapTop10) <-  "SDC"
```


```{r}
#save
##top10
Top10Ideotypes <- cbind(ScavengerTop10,SteepDeepCheapTop10)

colnames(Top10Ideotypes) <- c( "Topsoil foraging","Country", "Steep, Deep, Cheap","Country")

write.csv(Top10Ideotypes,"./tables/ideotypes/Top10Ideotypes_day18.csv", row.names = FALSE)

##top10
Bottom10Ideotypes <- cbind(ScavengerBottom10,SteepDeepCheapBottom10)

colnames(Bottom10Ideotypes) <- c( "Topsoil foraging","Country", "Steep, Deep, Cheap","Country")

write.csv(Bottom10Ideotypes,"./tables/ideotypes/Bottom10Ideotypes_day18.csv", row.names = FALSE)


##join using PI

ST <- Scavenger[,c(1,2,7), drop = FALSE]
DT <- SteepDeepCheap[,c(1,2,8), drop = FALSE]

AllIdeotypes <- merge(ST, DT) 
#grab the rankings of  other genotpes

write.csv(Scavenger, "./tables/iroot_ranks/foraging_Day18.csv", row.names = FALSE)

write.csv(SteepDeepCheap, "./tables/iroot_ranks/steepCheapDeap_Day18.csv", row.names = FALSE)

##Note this are how each genotype ranked in descending order. i.e Highest ranked first
write.csv(AllIdeotypes, "./tables/iroot_ranks/Day18_ideotypes_combined.csv", row.names = FALSE)

```

###############################end of day 18#################################################


############################################################################################
##################################### Day 15 ################################################
############################################################################################

```{r}
############################################################################################
#create a dataframe for ranking of each PI under each trait

#take only the columns with numerical data
colnum = c(5:(ncol(day15)-4)) # 

day15_ranks <- data.frame(matrix(vector(),235,1, dimnames=list(c(), c("PI")))) #create new dataframe that has 375 rows
day15_ranks$PI <- (day15$PI) #add column of PI numbers from dataframe

#colnames(day15)

i = 1
for (i in 1:length(colnum)) {  #this second loop runs through each TRAIT, one at a time
  x=colnum[i]  #set the current [i] column as x
  trait=colnames(day15)[x] #sets the current column header as the trait name
  #Rank <- as.data.frame(rank(desc(day15[x]))) #this was the original code. Needs to be relooked at...why desc. When merging, does it align with original PI/entry???
  Rank <- as.data.frame(rank(-(day15[x]))) #- means ranking from largest to smallest

    #Rank <- as.data.frame(rank(day15[x]))
  names(Rank)[1] <- trait  #set current trait as column header
  day15_ranks <- cbind(day15_ranks,Rank) #add columns to existing dataframe   
}

```
## iroot categories according to literature by Falk et al
## Each category is associated with certain traits
## will only focus on topsoil foraging and steep, deep and cheap

```{r}
############################################################################################
##################################### Scavenger ############################################
############################################################################################

IdeotypeVar <-  day15_ranks %>%
                            select(TRLUpper,WID, TRL_GR)
#select RSA traits to use for particular ideotype
########################


IdeotypeSums <- IdeotypeVar %>%
                            mutate(SUM = IdeotypeVar %>% rowSums()) #sum the rankings to make a master ranking column
Meta <- day15_ranks %>%
                  select(PI) #Create meta df with PI 

Scavenger <- cbind(Meta, IdeotypeSums) #bind meta and ideotype df together

#add country
country_1 <- day15[,c(1,3)] 
Scavenger <- merge(country_1, Scavenger, by="PI")
  
Scavenger <- arrange(Scavenger,SUM)  %>%
mutate(ScavengerRanked = rank(SUM, ties.method = "max")) #sort new df based on master ranking column with no.1 being best



ScavengerTop10 <- as.data.frame(Scavenger[1:10,1:2, drop = FALSE])
ScavengerBottom10 <- as.data.frame(Scavenger[225:235,1:2, drop = FALSE])

#colnames(ScavengerTop10) <-  "Scavenger"
```


```{r}
############################################################################################
################################## Steep, Deep, Cheap#######################################
############################################################################################

day15_ranks <- mutate(day15_ranks, LRADescending = 236 - LRA)  #invert the ranking for LRAMediangle as a lower angle, not higher angle is desirable

day15_ranks <- mutate(day15_ranks, SOL2 = -SOL2)  #invert the ranking for SOL2 as a lower density, not higher density is desirable

IdeotypeVar <-  day15_ranks %>%
                            select(LRADescending,PRL,SOL2,TRL_GR) #select RSA traits to use for particular ideotype
IdeotypeSums <- IdeotypeVar %>%
                              mutate(SUM = IdeotypeVar %>% rowSums()) #sum the rankings to make a master ranking column
Meta <- day15_ranks %>%
                   select(PI) #Create meta df with Entry, PI and Cluster info

SteepDeepCheap <- cbind(Meta, IdeotypeSums) #bind meta and ideotype df together

#add country
country_1 <- day15[,c(1,3)] 
SteepDeepCheap <- merge(country_1, SteepDeepCheap, by="PI")
 
SteepDeepCheap <- arrange(SteepDeepCheap, SUM) %>% mutate(DTRanked = rank(SUM, ties.method = "max")) #sort new df based on master ranking column with no.1 being best

SteepDeepCheapTop10 <- as.data.frame(SteepDeepCheap[1:10,1:2, drop = FALSE])
SteepDeepCheapBottom10 <- as.data.frame(SteepDeepCheap[225:235,1:2, drop = FALSE])
#colnames(SteepDeepCheapTop10) <-  "SDC"
```


```{r}
#save
##top10
Top10Ideotypes <- cbind(ScavengerTop10,SteepDeepCheapTop10)

colnames(Top10Ideotypes) <- c( "Topsoil foraging","Country", "Steep, Deep, Cheap","Country")

write.csv(Top10Ideotypes,"./tables/ideotypes/Top10Ideotypes_day15.csv", row.names = FALSE)

##top10
Bottom10Ideotypes <- cbind(ScavengerBottom10,SteepDeepCheapBottom10)

colnames(Bottom10Ideotypes) <- c( "Topsoil foraging","Country", "Steep, Deep, Cheap","Country")

write.csv(Bottom10Ideotypes,"./tables/ideotypes/Bottom10Ideotypes_day15.csv", row.names = FALSE)


##join using PI

ST <- Scavenger[,c(1,2,7), drop = FALSE]
DT <- SteepDeepCheap[,c(1,2,8), drop = FALSE]

AllIdeotypes <- merge(ST, DT) 
#grab the rankings of  other genotpes

write.csv(Scavenger, "./tables/iroot_ranks/foraging_Day15.csv", row.names = FALSE)

write.csv(SteepDeepCheap, "./tables/iroot_ranks/steepCheapDeap_Day15.csv", row.names = FALSE)

##Note this are how each genotype ranked in descending order. i.e Highest ranked first
write.csv(AllIdeotypes, "./tables/iroot_ranks/Day15_ideotypes_combined.csv", row.names = FALSE)

```

################################end of day15###############################################



############################################################################################
##################################### Day 12  ##############################################
############################################################################################


## iroot categories according to literature by Falk et al 2020
## Each category is associated with certain traits
## will only focus on topsoil foraging and steep, deep and cheap


###day12 cannot be used to rank scavenger because it does not use TRL_GR



####################################end of day12 ###########################################
